---
title: "Testing"
author: "E.M.Thomas"
date: "2023-02-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(here)
library(sf)
library(tidyverse)
library(terra) ### raster data functionality
library(tmap)
library(rgdal)
library(stars)
library(sp)
library(terra)
```




```{r}
land_cover_file <- here("nlcd_us", "nlcd_2019_land_cover_l48_20210604.img")

lc_rast <- terra::rast(land_cover_file)
lc_rast
```

```{r}
lc_1_rast <- lc_rast[[1]]
plot(lc_1_rast, main = 'US Landcover Map', 
     col = hcl.colors(20), axes = FALSE)
```


```{r}
cclip_file <- here("nlcd_us", "clipped_NY.tif")

cclip_rast <- terra::rast(cclip_file)
cclip_rast
```


```{r}
cclip_1_rast <- cclip_rast[[1]]
plot(cclip_1_rast, main = 'Chloe Clip Map', 
     col = hcl.colors(20), axes = FALSE)
```

  
```{r}
ny_vector <- read_sf(here("nlcd_us","NY_state.gpkg")) %>% 
  select(geom)
plot(ny_vector)
```

###attempting to rasterize NY vector 


```{r}
ny_rast_2 <- rasterize(ny_vector, aclip_rast)
```


```{r}
plot(ny_rast_2)
```




### making lc_rast a df
```{r}
us_df <- as.data.frame(lc_1_rast, xy = TRUE)
```

## trying to crop lc by ny 
```{r}
landtype_crop1 <- crop(lc_rast, ny_vector)

plot(landtype_crop, main = "Cropped lidar")
```

## trying to crop lc_rast by ny_vector
```{r}
landtype_crop <- crop(lc_rast, ny_rast_2)

plot(landtype_crop, main = "Cropped lidar")

```


```{r}
ggplot() +
  geom_raster(data = us_df, aes(x = x, y = y)) + 
  geom_sf(data = ny_vector, color = "blue", fill = NA) +
  coord_sf()
```

###dont run this 
```{r}
#plot_2<- plot(lc_1_rast, main = 'US Landcover Map', 
     #col = hcl.colors(20), axes = FALSE,
    # ny_vector, add = TRUE)
```

# add shapefile on top of the existing raster
plot(crop_extent, add = TRUE)




```{r}
hawaii_file <- here("nlcd_data", "hawaii_2001", "hi_landcover_wimperv_9-30-08_se5.img")

hawaii_rast <- terra::rast(hawaii_file)
hawaii_rast
```

```{r}
hawaii_1_rast <- hawaii_rast[[1]]
plot(hawaii_1_rast, main = 'hawaii Landcover Map', axes = FALSE)
```


```{r}
hawaii_parks_vector <- read_sf(here("nlcd_data","hawaii_2001", "parks_state","parks_state.shp")) 
hawaii_parks_vector
```

```{r}
crs(hawaii_rast)
```



```{r}
crs(hawaii_rast)
hawaii_x<- hawaii_parks_vector %>% 
  st_transform(st_crs(hawaii_rast))
```

```{r}
crs(hawaii_parks_vector)
```


```{r}
ak_park <- hawaii_parks_vector %>% 
   filter(NAME == "AHUPUAA O KAHANA STATE PARK") #%>% 
 # select(geometry)
plot(ak_park)
```



```{r}
plot(hawaii_1_rast, main = 'hawaii Landcover Map', axes = FALSE) 
plot(hawaii_parks_vector %>% 
       select(geometry), add = TRUE)
```


```{r}
hawaii_coarse <- terra:: aggregate(x= hawaii_rast, fact =4, fun = modal)
```


```{r}
writeRaster(hawaii_coarse, filename = "lc_rast_coarse.tif")
```

```{r}
lc_coarse_file <- here("nlcd_data", "lc_rast_coarse.tif")

lc_rast <- terra::rast(lc_coarse_file)
lc_rast
```


```{r}
hawaii_df <- as.data.frame(hawaii_coarse, xy = TRUE)
```

### Integrating the Hawaii df (containing pixel and class data) with carbon calculation + carbon map:

```{r}
hawaii_class_areas <- hawaii_df %>% 
  janitor::clean_names() %>% 
  group_by(land_cover_class) %>% 
  tally(name = "area_hectares") ### Counts the number of observations per group (defined as land cover class in previous row).
```

```{r}
ratio_carbon_tph <- readxl::read_xlsx(here("carbon_stock_cfs/carbon_storage_ton_C_per_hectare.xlsx")) %>% 
  clean_names() %>% 
  dplyr::select(class:soc) %>% 
  pivot_longer(cols = above:soc, 
               names_to = "compartment", values_to = "ton_c_per_hectare")

hawaii_total_carbon <- merge(hawaii_class_areas, ratio_carbon_tph, 
                           by.x = "land_cover_class", by.y = "class") %>% 
  mutate(total_carbon_tons_log = log(area_hectares * ton_c_per_hectare)) 

hawaii_carbon_plot <- ggplot() +
  geom_col(data = hawaii_total_carbon, 
           mapping = aes(x = biome, 
                         y = total_carbon_tons_log, 
                         fill = compartment
                         )) +
  scale_fill_manual(values = c("plum2", "slateblue1", "palegreen1", "lightgoldenrod1", "tan1"), 
                    labels = c(above = "Above Ground", 
                               below = "Below Ground", 
                               dead_matter = "Dead Matter", 
                               litter = "Litter", 
                               soc = "Soil Organic Carbon"
                               )) +
  theme(axis.text.x = element_text(angle = 45, 
                                   vjust = 1, 
                                   hjust=1
                                   )) +
  labs(x = "", 
       y = "Carbon (log tons)", 
       title = "Carbon Storage Per Land Use Type in Hawaii", 
       fill = "Compartment")

hawaii_carbon_plot
```

### Further plots by Chloe

``` {r}
ggplot() +
  geom_raster(data = hawaii_df, aes(x = x, y = y, fill = ndvi)) +
  coord_sf(expand = 0) +
  scale_fill_gradient(low = 'grey30', high = 'grey80') +
  theme_void() +
  theme(panel.background = element_rect(fill = 'slateblue4'))
```


```{r}
ggplot() 
  geom_raster(data = hawaii_df, aes(x = x, y = y)) +
  coord_sf(expand = 0, aes(color = "Land Cover Class")) + 
  theme_void() 
```
crop is square result 
mask is the exact shape of the polygon 



```{r}
unique(hawaii_df$`Land Cover Class`)
```

labels = c("NA", "Open Water", "Developed, Open Space", "Developed, Low Intensity", "Developed, Medium Intensity", "Developed, High Intensity", "Barren Land (Rock/Sand/Clay)", "Evergreen Forest", "Shrub/Scrub", "Grassland/Herbaceous", "Pasture/Hay", "Cultivated Crops", "Woody Wetlands", "Emergent Herbaceous Wetlands")

```{r}
tmap_mode("view")
#tmap_options(check.and.fix = TRUE)
tm_shape(lc_rast) +
  tm_raster(style = "cat", palette = c("lightblue", "blue", "lightpink", "coral1", "red", "darkred", "tan", "darkgreen", "darkgoldenrod3", "darkkhaki", "khaki1", "brown", "lightcyan", "lightseagreen"), labels = c("NA", "Open Water", "Developed, Open Space", "Developed, Low Intensity", "Developed, Medium Intensity", "Developed, High Intensity", "Barren Land (Rock/Sand/Clay)", "Evergreen Forest", "Shrub/Scrub", "Grassland/Herbaceous", "Pasture/Hay", "Cultivated Crops", "Woody Wetlands", "Emergent Herbaceous Wetlands"), n= 14, title = "Land Type") +
  tm_shape(hawaii_x) +
  tm_borders(col = "black", lwd = 2) 
```



```{r}
unique(hawaii_df$`Land Cover Class`)
```



palette = c("blue", "green", "darkgreen", "orange", "yellow")
```{r}
tmap_mode("view")

tm_shape(hawaii_coarse) +
  tm_raster(palette = c(
    "0" = "lightblue",
    "11" = "blue",
    "31" = "pink",
    "22" = "red",
    "52" = "green",
    "42" = "darkgreen"
  ), n= 14)
```



```{r}
roi_vec <- read_sf(here("nlcd_data",
                        "hawaii_2001", 
                        "parks_state",
                        "parks_state.shp")) %>% 
  st_transform(st_crs(hawaii_rast))

crs(roi_vec)
plot(roi_vec)
```


### Eleanor trying stuff

```{r}
# Read in shapefile
hawaii_parks_vector <- read_sf(here("nlcd_data","hawaii_2001", "parks_state","parks_state.shp")) 

# Filter for specific national park 
ak_park <- hawaii_parks_vector %>% 
   filter(NAME == "AHUPUAA O KAHANA STATE PARK") #this turns ak_park into an sp file because we used the filter function

```


```{r}
# Define region of interest vector 
roi_sf <- ak_park

# Project vector file to match raster file
roi_sf_proj <- st_transform(roi_sf, crs(lc_rast))

# Crop raster 
lc_roi_rast <- crop(lc_rast, roi_sf_proj)

# Mask raster with vector file
lc_roi_masked_rast <- mask(lc_roi_rast, roi_sf_proj)
plot(lc_roi_masked_rast)

```

